I want to first give thanks to my advisor, Jeroen Tromp.
It has been one of my best experiences to work with him for the past six years.
The more I worked with him, the more I realized he is such a talented,
knowledgeable and easy-going person. He is also willing to help his students anytime.
Whenever I had a question, I just went and knocked on his door. He would always do his best to help me.
If it was a questions related to my research, he would start from the very basics to
answer my question in great detail.
There were times when we encountered difficulties and challenges during the past six years.
Jeroen also has the personality to encourage us and go back to the right track.
He has great vision and is always dedicated to achieving the overall goal, and he is willing to
share his optimism and courage with others.

I also want to give thanks to my committee members: Allan Rubin, Frederik Simons,
and Jessica Irving. They provided me a lot of help during my generals and let me
pass it safely. In my global projects, they also provided numerous suggestions
and insights that help me to better understand the model and create new ideas.

I would like to acknowledge my colleagues in the seismology group: Youyi Ruan, Hom Nath Gharti, Matthieu Lefebvre, James Smith, Ryan Modrak, Yanhua Yuan, Uno B.~Vaaland, David Luet...
I have been working Youyi extensively on the global project for last 3 years. He is one of the best collaborators and we worked so numerous times together to tackle problems. 
Hom Nath Gharti has been my office mate for four years.
He is such a knowledgeable and friendly person. It is one of my best memories to have stayed in Room 309, and I feel that was my ``home''.
It is hard to image not having Matthieu in
our group for the past six years. I learned so much technical knowledge from him. He was always the first person I would go to to ask for help if I had a technical
problem.
It was also great to work with James Smith and Ryan Modrak on the global
project and many associated sub-projects. They were the persons I could always
rely on.

I would also like to take this opportunity to acknowledge several previous members of our
group: Ebru Bozda\u g, Hejun Zhu, Qinya Liu, Min Chen, Carl Tape, Dimitri Komatitsch
Daniel Peter... Their work built the foundation of the adjoint tomography workflow.
I also want to give special thanks to Ebru, Hejun, and Min. They provided me with
their models, so I can compare my models with theirs. I also had numerous discussions
with them at AGU meetings and on other occasions, which provided me with new insights into
our model, bringing GLAD-M25 to fruition. Just before the submission of our paper detailing
GLAD-M25, I was shocked to hear that Dimitri Komatitsch passed away unexpectedly.
He is a extraordinary computational geophysicist who built the family of SPECFEM
software together with Jeroen.

I want to thank students and friends in the Geosciences Department: Yajun Peng, Wenbo Wu, 
Pathikrit Bhattacharya, Farhan Nuruzzaman, and Cara Magnabosco. It was always joyful to talk
with you. I give my best wishes to our administrative assistants: Mary
Rose Russo and Sheryl Robas. Their hard work and efforts make it possible for us
to focus on our research.

I want to thank my former advisor at USTC, Professor Xiaofei Chen. I did a summer project with him
and he introduced me to geophysics. From him and Jeroen, I feel their passion
for and dedication to geophysical research. They are my role models in my future career.

I give final thanks to my parents, Jinfu Lei and Shenglan Lou, and my girlfriend Qizhi Xu.
It is your support, love, and care that make my life so meaningful and joyful, and make
me become who I am today.

